
\section{Week 13}

\subsection{Report}

Rewrote the introduction and the portion of the report on conception of the idea.  The impetus was largely a lot of propaganda—the clear demonstration that people are not sufficiently educated to do more than follow along with whatever they're told by a convincing narrator.  FairVote in particular has been driving the acceptance of IRV and rejection of Condorcet methods; their rhetoric even came up at the Society of Public Choice's annual member's meeting, criticizing the inconsistencies and the shoddy data used.

\subsection{Lessons}

Expanded the overview of \hyperref[cur:ed-history]{Education and History of Social Choice}.

Created the overview for \hyperref[cur:public-discourse]{Public Discourse}.and \hyperref[cur:manipulation]{Manipulation}.

\hyperref[cur:representative-democracy]{Representative Democracy}, \hyperref[cur:evaluating-decision-methods]{Evaluating Decision Methods}, and \hyperref[cur:government-structure]{Government Structure} still need writing.

Started adding notes for good places to talk about slavery, women's rights, felony disenfranhcisement, and immigration and citizenship.

Enumerated traps for several lessons.  Certain discussions and questions are prone to reaching incorrect conclusions; these are listed where they're easy to bait.

I put \hyperref[cur:evaluating-decision-methods]{Evaluating Decision Methods} after \hyperref[cur:government-structure]{Government Structure}.  It has little value in the voting rules and government structure lessons; but there's nothing to teach during that lesson, whereas the concepts provide a \textit{huge} amount of potential for debate and discourse tying together all the prior lessons.

\subsection{Readings}

The readings now render as separate documents.  A preamble reconfiguring the section headings removes the appendix title from the documents.

I added several voting rules to the appendix on voting rules.

Found out Collective Decisions and Voting \autocite{Tideman2006} is available as an ebook from interlibrary loan, so I've been assiging readings out of that.

\subsection{Ideas}

Not sure if it's worth taking a day to discuss FairVote, the range voting site, and the center for election innovation and researhc as examples of abuse of people's lack of knowledge via propaganda; National Popular Vote for efforts undertaken from a lack of knowledge; and other such issues.  This would bring everything together to create perspective.
