\chapter{Electoral Manipulation}

\section{Summary}

\subsection{Content}


\subsection{Annotations}

\section{Sources}


\section{Concepts}

\section{Missing Information}

\subsection{Manipulating Final-Five}

Final Five uses SNTV as the primary election, and IRV as the general election.  A few facts to research and tools to use:

\begin{itemize}
    \item Manipulating the single non-transferable vote primary
    \begin{itemize}
        \item Japan had trouble with SNTV being manipulated for years.

        \item Figure out how many marginal opponents can be weakened to drive votes to unelectable alternatives, or to strong opponents who cannot win.

        \item Support canvassing efforts for the opposition party in areas with the strongest demographic support for these marginal opponents.

        \item Have agents engage the campaigns of weak opposition party candidates, providing strategy effective against the most centrist candidate—the one most likely to beat yours in the general election.

        \item Target campaign efforts to different sets of party voters, targeting a minimum number of votes above the figure you expect your weakest opponent to get \textit{after} efforts to weaken them.

        \item Minimize your own party's excess candidates, and use demographics research to figure model a rough STV outcome among the party to determine which candidates to support.
    \end{itemize}

    \item Manipulating Instant Runoff Voting
    \begin{itemize}
        \item The normal approaches to campaign-based attacks work.

        \item If possible, influence which of the opposition candidates are ultimately nominated in the primary, aiming to get the more centrist candidate eliminated in the general election.  This shifts swing voters onto your party's side.

        \item If one of your underdogs displaced the opposition's centrist candidate, reduce support or have your agents disrupt their campaign.
    \end{itemize}
\end{itemize}

When considering the median voter as favoring the opposition party, your goal is to get nominated a stronger centrist candidate of your party, and to control the opposition nominations such that the more centrist opponent is eliminated before yours.  This moves the opposition party's candidate farther away from the median voter, which nets your party more votes and can turn out a win.  The extra votes from swing voters will pad your candidate when the opponent's centrist is eliminated, further protecting them from elimination.

Eliminating the centrist in the primary and leaving a more extreme candidate can provide the same outcome, and relies on stronger opposition party turn-out weighted further from the median voter, or else on dividing the opposition's most centrist votes among multiple candidates.

Negative campaigning and well-timed manufactured scandal can have an effect.  Watchdog groups will figure out when one party is running ads supporting another party's candidate in the primary, but most people are not reading the feed.